---
title: "Statistical Inference: The connection between the exponential distribution and the central limit theorem"
author: "Marc T. Henry de Frahan"
output: 
  html_document:
    keep_md: true
---
# Synopsis

# Load some libraries
```{r}
library(ggplot2)
```


# Setup

For the rest of this project, we set the rate parameter, lambda, to
0.2, the number of simulations, nsims, to 1000, and the number of drawings for each simulation, ndraw, to 40.

```{r}
lambda <- 0.2
mean_theory <- 1/lambda
nsims <- 1000
ndraw <- 40
```

Let's run `r nsims` simulations of the exponential distribution. Each
time we do a simulation, we save the mean to a vector of means.

```{r}
sim_means = NULL
for (i in 1 : 1000) sim_means = c(sim_means, mean(rexp(ndraw,lambda)))
```

# Comparing the sample mean to the theoretical mean

The sample mean is `r mean_sim` and the theoretical mean is `r
mean_theory`. As you can see in this figure, the sample distribution
is very closely centered to the theoretical distribution.

```{r} histogram of simulations
ggplot() +
    aes(sim_means) +
    geom_histogram(binwidth=0.1) +
    geom_vline(aes(xintercept=mean_theory, colour="Mean.Th", linetype = "Mean.Th"), size = 1.5) +
    geom_vline(aes(xintercept=mean_sim, colour="Mean.Sim", linetype = "Mean.Sim"), size = 1.5, show.legend=TRUE) +
    scale_colour_manual(name="Labels", values=c(Mean.Sim="#009E73",Mean.Th="black")) +
    scale_linetype_manual(name="Labels", values=c(Mean.Sim="dashed",Mean.Th="solid"), guide=FALSE)+
    labs(x = "Value",
         y = "Counts",
         title = "Histogram of exponential distribution simulations")
```


# Comparing the sample variance to the theoretical variance

