---
title: "Statistical Inference: The connection between the exponential distribution and the central limit theorem"
author: "Marc T. Henry de Frahan"
output: 
  html_document:
    keep_md: true
---
# Synopsis

# Load some libraries
```{r}
library(ggplot2)
```


# Setup

We set lambda, the rate parameter, to 0.2 for the rest of this project and the number of simulations, n, to 1000.
```{r}
lambda <- 0.2
mean_theory <- 1/lambda
n <- 1000
```

Let's run `r n` simulations of the exponential distribution and save
them to a variable. We also compute the sample mean.
```{r}
sim <- rexp(n, lambda)
mean_sim <- mean(sim)
```

# Comparing the sample mean to the theoretical mean

The sample mean is `r mean_sim` and the theoretical mean is `r
mean_theory`. As you can see in this figure, the distribution is very
closely centered next to the theoretical distribution.

```{r} histogram of simulations
ggplot() +
    aes(sim) +
    geom_histogram(binwidth=1) +
    geom_vline(aes(xintercept=mean_theory, colour="Mean.Th", linetype = "Mean.Th"), size = 1.5) +
    geom_vline(aes(xintercept=mean_sim, colour="Mean.Sim", linetype = "Mean.Sim"), size = 1.5, show.legend=TRUE) +
    scale_colour_manual(name="Labels", values=c(Mean.Sim="#009E73",Mean.Th="black")) +
    scale_linetype_manual(name="Labels", values=c(Mean.Sim="dashed",Mean.Th="solid"), guide=FALSE)+
    labs(x = "Value",
         y = "Counts",
         title = "Histogram of exponential distribution simulations")
```

